Object-oriented implementations of the MPDATA advection equation solver in~C++,~Python~and~Fortran

### 2013-05-02 paper/doc/paper.tex
hint on the CMake 2.8.10 bug

  \begin{figure}[h!]
  \center
  \begin{tikzpicture}
    \coordinate (Origin)   at (0,0);
    \coordinate (XAxisMin) at (-2.5,0);
    \coordinate (XAxisMax) at (3,0);
    \coordinate (YAxisMin) at (0,-2.5);
    \coordinate (YAxisMax) at (0,3);
    \draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
    \draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis

    \clip (-2.5,-2.5) rectangle (2.5cm,2.5cm); % Clips the picture...
    \draw[style=help lines,dashed] (-14,-14) grid[step=2cm] (14,14);
          % Draws a grid in the new coordinates.
          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
              % Puts the shaded rectangle
    \foreach \x in {-7,-6,...,7}{% Two indices running over each
      \foreach \y in {-7,-6,...,7}{% node on the grid we have drawn 
        \node[draw,circle,inner sep=2pt,fill] at (2*\x,2*\y) {};
            % Places a dot at those points
      }
    }
    \draw [black] (0,0) -- (0,0) node [below right] {$\!\psi_{[i,j]}$};
    \draw [black] (-2,0) -- (-2,0) node [below right] {$\!\psi_{[i-1,j]}$};
    \draw [black] (0,2) -- (0,2) node [below right] {$\!\psi_{[i,j+1]}$};

    \draw [ultra thick,-latex,red] (.8,0) -- (1.2,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i+\phlf,j]}$};
    \draw [ultra thick,-latex,red] (-1.2,0) -- (-.8,0) node [above right] {$\!\!\!\!\!\!\!\!\!\!C^{[x]}_{[i\mhlf,j]}$};
    \draw [ultra thick,-latex,red] (.8,2) -- (1.2,2) node [above] {};
    \draw [ultra thick,-latex,red] (-1.2,2) -- (-.8,2) node [above] {};
    \draw [ultra thick,-latex,red] (.8,-2) -- (1.2,-2) node [above] {};
    \draw [ultra thick,-latex,red] (-1.2,-2) -- (-.8,-2) node [above] {};

    \draw [ultra thick,-latex,red] (0,.8) -- (0,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (0,-1.2) -- (0,-.8) node [below right] {$C^{[y]}_{[i,j\mhlf]}$};
    \draw [ultra thick,-latex,red] (-2,.8) -- (-2,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (-2,-1.2) -- (-2,-.8) node [above] {};
    \draw [ultra thick,-latex,red] (2,.8) -- (2,1.2) node [above] {};
    \draw [ultra thick,-latex,red] (2,-1.2) -- (2,-.8) node [above] {};
  \end{tikzpicture}
  \caption{\label{fig:grid}
    A schematic of the Arakawa-C grid.
  }
  \end{figure}
  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}
    is a natural choice for MPDATA.
  As a consequence, the discretised representations of the $\psi$ scalar field, and each 
    component of the ${\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}}$ vector field 
    in eq.~(\ref{eq:adv}) are defined over different grid point locations.
  In mathematical notation this can be indicated by usage of fractional indices, e.g.
    $C^{[x]}_{[i\mhlf,j]}$, $C^{[x]}_{[i+\phlf,j]}$, $C^{[y]}_{[i,j\mhlf]}$ and $C^{[y]}_{[i,j+\phlf]}$
    to depict the grid values of the ${\vec{C}}$ vector components surrounding $\psi_{[i,j]}$.
  However, fractional indexing does not have a built-in counterpart in any of the
    employed programming languages.
  A desired syntax would translate \prog{$i-\phlf$} to \prog{$i-1$} and
    \prog{$i+\phlf$} to \prog{$i$}.
  OOP offers a convenient way to implement such notation
    by overloading the \prog{+} and \prog{-} operators for objects representing array indices.

@@

  \begin{figure}[t]
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{laptok-cpu}
    \caption{\label{fig:cpu-eyrie}
      Execution time statistics for the test runs described in Section~\ref{sec:perf}
        plotted as a function of grid size.
      Values of the total user mode CPU time reported by the GNU time utility are
        normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
      Before normalisation the time reported for an $nt=0$ run for a corresponding
        domain size is subtracted from the values.
      Both the $nt=0$ and $nt=2^{24}/()nx \cdot ny$ runs are repeated three times and
        only the shortest time is taken into account.
      Results obtained with an Intel\textsuperscript{\textregistered} 
        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
    }
  \end{figure}
  \begin{figure}[t]
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{skua-cpu}
    \caption{\label{fig:cpu-skua}
      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
    }
  \end{figure}

>>>>

  \begin{figure}[h!]
  \center
  \input{grid}
  \caption{\label{fig:grid}
    A schematic of the Arakawa-C grid.
  }
  \end{figure}
  The so-called Arakawa-C staggered grid \citep{Arakawa_and_Lamb_1977} depicted in Figure~\ref{fig:grid}
    is a natural choice for MPDATA.
  As a consequence, the discretised representations of the $\psi$ scalar field, and each 
    component of the ${\vec{C}=\vec{v}\cdot\frac{\Delta t}{\Delta x}}$ vector field 
    in eq.~(\ref{eq:adv}) are defined over different grid point locations.
  In mathematical notation this can be indicated by usage of fractional indices, e.g.
    $C^{[x]}_{[i\mhlf,j]}$, $C^{[x]}_{[i+\phlf,j]}$, $C^{[y]}_{[i,j\mhlf]}$ and $C^{[y]}_{[i,j+\phlf]}$
    to depict the grid values of the ${\vec{C}}$ vector components surrounding $\psi_{[i,j]}$.
  However, fractional indexing does not have a built-in counterpart in any of the
    employed programming languages.
  A desired syntax would translate \prog{$i-\phlf$} to \prog{$i-1$} and
    \prog{$i+\phlf$} to \prog{$i$}.
  OOP offers a convenient way to implement such notation
    by overloading the \prog{+} and \prog{-} operators for objects representing array indices.

@@

  \begin{figure}[t]
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{laptok-cpu}
    \caption{\label{fig:cpu-eyrie}
      Execution time statistics for the test runs described in Section~\ref{sec:perf}
        plotted as a function of grid size.
      Values of the total user mode CPU time reported by the GNU time utility are
        normalised by the grid size ($nx \cdot ny$) and the number of timesteps $nt=2^{24}/(nx \cdot ny)$.
      Before normalisation the time reported for an $nt=0$ run for a corresponding
        domain size is subtracted from the values.
      Both the $nt=0$ and $nt=2^{24}/(nx \cdot ny)$ runs are repeated three times and
        only the shortest time is taken into account.
      Results obtained with an Intel\textsuperscript{\textregistered} 
        Core\textsuperscript{\texttrademark} i5 1.6 GHz processor.
    }
  \end{figure}
  \begin{figure}[t]
    \center
    \includegraphics[height=.45\textwidth,angle=-90]{skua-cpu}
    \caption{\label{fig:cpu-skua}
      Same as Fig.~\ref{fig:cpu-eyrie} for an AMD Phenom\textsuperscript{\texttrademark} II 800 MHz processor.
    }
  \end{figure}
